|

### 2025-11-27 manuscript/submission_manuscript.qmd
preprint version formatted, submission version to finish

        \usepackage{setspace}\doublespacing

@@

        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}

@@

  Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age women, who often have higher BMI classification. This is assumed to stem from lower resting energy expenditure (REE), influencing lifestyle intervention guidelines. However, evidence for reduced REE in women with PCOS compared with those without is inconsistent. Objective: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS. Data Sources: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025. Study Selection: Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included. Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate. Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors was used to compare both mean REE, and between person variation in REE, between women with and without PCOS. Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33]. Conclusions: These findings indicate that REE does not meaningfully differ between women with and without PCOS. Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.

@@

# Materials and Methods
This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@pagePRISMA2020Statement2021a]. The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?". We summarise and describe the studies in addition to quantitatively synthesising their results via meta-analysis.

@@

@@

Note, the numbers in this section exclude the four studies previously noted with possible reporting issues that were unresolved. These studies are however summarised in @tbl-review where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al. [-@saltamavrosAlpha2Beta2007].

Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`). Descriptive characteristics of the arms and participants in these studies are reported in @tbl-review.

@@

Studies were carried out in multiple countries: Brazil (k = `r filter(countries, country == "Brazil")$n`) ,and USA (k = `r filter(countries, country == "USA")$n` studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK (all k = 1 study). The four studies with noted reporting errors were carried out in Greece. A total of 11 studies used the Rotterdam criteria (or a modified version there of) for diagnosing PCOS [@robinsonPostprandialThermogenesisReduced1992; @brunerEffectsExerciseNutritional2006; @moranShorttermMealReplacements2006; @cosarRestingMetabolicRate2008; @graffDietaryGlycemicIndex2013; @pohlmeierEffectLowstarchLowdairy2014; @dohRelationshipAdiposityInsulin2016; @larssonDietaryIntakeResting2016; @graffSaturatedFatIntake2017; @rodriguesLowValidityPredictive2018; @tosiInsulinMediatedSubstrateUse2021]. One study used the 1990 National Institutes of Health criteria [@broskeyAssessingEnergyRequirements2017], and two studies [@segalRestingMetabolicRate1990; @robinsonPostprandialThermogenesisReduced1992] diagnosed PCOS via the presence of oligomenorrhea or amenorrheaammenorrhea alongside additional criteria including plasma androgen levels, hirsutism or polycystic ovaries on ultrasound scanning. All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry. The specific devices reported by these studies are included in @tbl-review. We originally considered in our pre-registration that, given sufficient data, we would compare sub groups of women with PCOS who did and did not have accompanying insulin resistance. However, based on the metabolic health variables reported in studies with this information (see @tbl-review) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance. Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging from `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`. Across those studies where BMI was reported or it was possible to estimate the mean BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]`.

![Descriptive characteristics of arms and participants for included studies](descriptives_table.png){#tbl-review}

@@

```{r}
#| label: fig-mean-REE 
#| fig-width: 10
#| fig-height: 7.5
#| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.

@@

```{r}
#| label: fig-variance-REE 
#| fig-width: 10
#| fig-height: 7.5
#| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.

@@

@@

This study sought to estimate and describe the magnitude of difference in REE between women with and without PCOS. Most studies identified in the systematic review, and included in the meta-analysis, used indirect calorimetry as the primary measure of REE and assessed women with PCOS who were insulin resistant and categorised as being in overweight or obese BMI categories compared to healthy controls. Our results indicate there is only a small magnitude of difference in REE (`r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)`kcal/day]) between women with PCOS and those without. Further, there is little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95%QI: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]) suggesting that, despite individual differences in REE, PCOS is not systematically associated with lesser or greater individual variability. These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain. This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021]. Recognising that there may be minimal differences in REE between women with PCOS and those without can inform both clinical and public practices, potentially leading to a shift in focus away from a requirement for more severe caloric restriction as a primary method of treatment towards more comprehensive, individualised, and psychologically safe approaches to care [@teedeRecommendations2023International2023].

```{r}
targets::tar_load(c(main_arm_mean_effects_contrast_condition_bmi, main_arm_mean_effects_contrast_condition_fat_free_mass))

The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively). If anything, REE may be slightly higher in women with PCOS compared to healthy women without PCOS. BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]` and this may explain the slightly greater REE in the former group. However, for those studies where we could extract or estimate BMI, we fit an additional exploratory model extending the main arm based model and extracted an estimate of the difference adjusted for BMI as a fixed effect. Adjusted to the median value of average BMI of the control arms of women without PCOS (BMI = `r main_arm_mean_effects_contrast_condition_bmi$m_bmi` kg/$\textrm{m}^2$) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)`kcal/day]) between women with PCOS and those without. However, one of the studies included in our analysis [@tosiRestingEnergyExpenditure2024] reported that, whilst there was little difference in unadjusted REE, when REE was adjusted for fat-free mass it was lower in women with PCOS suggesting the potential importance of fat-free tissue in energy regulation. Yet, similarly to the additional exploratory model including BMI, when we adjusted for fat-free mass including those studies reporting this characteristic (fat-free mass = `r main_arm_mean_effects_contrast_condition_fat_free_mass$m_fat_free_mass` kg) the difference in REE between was (`r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)`kcal/day]). As such, even adjusted for both BMI and fat-free mass, there seems to be little difference in REE between women with, and without, PCOS. Yet, a recent systematic review of mechanisms for metabolic dysfunction has reported excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780). Taken together, these findings highlight that factors beyond REE and typical correlates of this including BMI or fat-free mass, such as hormonal and tissue-specific metabolic effects, may play a more significant role in weight regulation challenges in women with PCOS.

>>>>

        \usepackage{setspace}\AtBeginDocument{\spacing{2}}

@@

        \usepackage{pdflscape}

@@

  Context: Polycystic ovary syndrome (PCOS) is common in reproductive-age women, who often have higher BMI classification. This is assumed to stem from lower resting energy expenditure (REE), influencing lifestyle intervention guidelines. However, evidence for reduced REE in women with PCOS compared with those without is inconsistent. Objective: To systematically search and meta-analyse the existing literature to estimate and describe the difference in REE between women with and without PCOS. Data Sources: A systematic search was conducted using PubMed, Medline and Web of Science databases of published research from January 1990 to January 2025. Study Selection: Studies that measured REE in women living with PCOS, both with and without control arms of women without PCOS, were included. Data Extraction: Bibliometric, demographic, and REE data was extracted by one investigator and checked in triplicate. Data Synthesis: Thirteen studies were included in a Bayesian arm-based multiple condition comparison (i.e., network) type meta-analysis model with informative priors to compare both mean REE, and between person variation in REE, between women with and without PCOS. Mean REE differed between groups by 30 kcal/day [95% quantile interval: -47 to 113 kcal/day] and the contrast ratio for between person standard deviations was 0.98 [95% quantile interval: 0.71 to 1.33]. Conclusions: These findings indicate that REE does not meaningfully differ between women with and without PCOS. Group-level differences in resting energy expenditure are small, insignificant, or not physiologically relevant.

@@

# Materials and Methods
This systematic review and meta-analysis was pre-registered on PROSPERO ([CRD42024601434](https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434)) initially on the $3^{rd}$ of December 2024 and performed in accordance with the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement guidelines [@pagePRISMA2020Statement2021a]. The PRISMA flow diagram reported below (@fig-prisma) was produced using the `PRISMA2020` R package and Shiny app [@haddawayPRISMA2020PackageShiny2022]. The primary aim of this review was to examine the descriptive question "Does resting energy expenditure (REE) differ between women with and without polycystic ovary syndrome (PCOS)?". We summarise and describe the studies in addition to quantitatively synthesising their results via meta-analysis.

@@

```{r}
targets::tar_load(c(main_arm_mean_effects_contrast_condition_bmi, main_arm_variance_effects_contrast_condition_bmi,
                    main_arm_mean_effects_contrast_condition_fat_free_mass,main_arm_variance_effects_contrast_condition_fat_free_mass))

Lastly, given the role of body mass and, in particular, highly metabolically active tissue on REE we included models adjusted for these group level characteristics where studies reported them. These models were essentially extensions of the main models noted above for BMI and fat-free mass where each was modelled with it's corresponding sampling error similarly to how the log mean REE was modelled in the models for standard deviation of REE i.e., the mean BMI or fat-free mass estimates were modelled as estimated with measurement error. When extracting posterior distributions for the contrasts between conditions in these models both BMI and fat-free mass where adjusted to the median values seen in the control conditions i.e., BMI = `r main_arm_mean_effects_contrast_condition_bmi$m_bmi` kg/$\textrm{m}^2$ and fat-free mass = `r main_arm_mean_effects_contrast_condition_fat_free_mass$m_fat_free_mass` kg.

@@

Note, the numbers in this section exclude the four studies previously noted with possible reporting issues that were unresolved. These studies are however summarised in fully in the descriptive characteristics table in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html) where some of the possible reporting errors are seen in the standard deviations reported or calculated e.g., the standard deviation calculated for body mass of women with PCOS in Saltamavros et al. [-@saltamavrosAlpha2Beta2007]. The PRISMA flow diagram (@fig-prisma) also includes all studies identified including the four with possible errors.

![PRISMA 2020 flow diagram template for systematic reviews. Note that a “report” could be a journal article, preprint, conference abstract, study register entry, clinical study report, dissertation, unpublished manuscript, government report or any other document providing relevant information. The website noted was a prior narrative review on this topic by some of the authors (https://macrofactorapp.com/pcos-bmr/) which identified two studies not found in our systematic database search.](prisma.png){#fig-prisma}

Our systematic review identified `r length(unique(main_arm_data_imputed_demographics$study))` studies from `r length(unique(main_arm_data_imputed_demographics$lab))` lab/research groups including `r length(unique(main_arm_data_imputed_demographics$arm))` arms (Control arms = `r no_conditions$n[1]`, PCOS arms = `r no_conditions$n[2]`) and a total of `r sum(no_participants$total_n)` participants (Controls: minimum n = `r no_participants$min_n[1]`, median n = `r no_participants$median_n[1]`, maximum n = `r no_participants$max_n[1]`; PCOS: minimum n = `r no_participants$min_n[2]`, median n = `r no_participants$median_n[2]`, maximum n = `r no_participants$max_n[2]`). Descriptive characteristics of the arms and participants in these studies are reported fully in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)).

@@

Studies were carried out in multiple countries: Brazil (k = `r filter(countries, country == "Brazil")$n`) ,and USA (k = `r filter(countries, country == "USA")$n` studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK (all k = 1 study). The four studies with noted reporting errors were carried out in Greece. A total of 11 studies used the Rotterdam criteria (or a modified version there of) for diagnosing PCOS [@robinsonPostprandialThermogenesisReduced1992; @brunerEffectsExerciseNutritional2006; @moranShorttermMealReplacements2006; @cosarRestingMetabolicRate2008; @graffDietaryGlycemicIndex2013; @pohlmeierEffectLowstarchLowdairy2014; @dohRelationshipAdiposityInsulin2016; @larssonDietaryIntakeResting2016; @graffSaturatedFatIntake2017; @rodriguesLowValidityPredictive2018; @tosiInsulinMediatedSubstrateUse2021]. One study used the 1990 National Institutes of Health criteria [@broskeyAssessingEnergyRequirements2017], and two studies [@segalRestingMetabolicRate1990; @robinsonPostprandialThermogenesisReduced1992] diagnosed PCOS via the presence of oligomenorrhea or amenorrheaammenorrhea alongside additional criteria including plasma androgen levels, hirsutism or polycystic ovaries on ultrasound scanning. All but one study [@broskeyAssessingEnergyRequirements2017], which used doubly labelled water, measured REE using indirect calorimetry. The specific devices reported by these studies are included in the online supplementary materials (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)). We originally considered in our pre-registration that, given sufficient data, we would compare sub groups of women with PCOS who did and did not have accompanying insulin resistance. However, based on the metabolic health variables reported in studies with this information (see [Descriptives Table](https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html)) and, where available, considering primary criteria of either homeostatic model assessment of insulin resistance (HOMA-IR) $\geq2.5$ or secondary criteria including fasting insulin $>12\mu \textrm{U/mL}$ or fasting glucose $\geq100\textrm{mg/dL}$, all groups of women with PCOS in the included studies would be considered to have insulin resistance. Mean age of women with PCOS in these studies ranged from `r summary_descriptives$min_age[2]` to `r summary_descriptives$max_age[2]` which was similar to the control women without PCOS ranging from `r summary_descriptives$min_age[1]` to `r summary_descriptives$max_age[1]`. Across those studies where BMI was reported or it was possible to estimate the mean BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]`.

@@

```{r}
#| label: fig-mean-REE 
#| fig-width: 10
#| fig-height: 6.25
#| fig-cap: Posterior distribution, mean point estimates, and 95% quantile intervals for conditional estimates by study, global grand mean estimates by condition, and the contrast between conditions for mean resting energy expenditure of control women without PCOS and women with PCOS.

@@

```{r}
#| label: fig-variance-REE 
#| fig-width: 10
#| fig-height: 6.25
#| fig-cap: Posterior distribution, variance point estimates, and 95% quantile intervals for conditional estimates by study, global grand variance estimates by condition, and the contrast ratio between conditions for the standard deviation of resting energy expenditure of control women without PCOS and women with PCOS.

@@

### BMI and Fat-Free Mass Adjusted Models
For mean REE the model adjusted for BMI, also resulted in qualitatively similar inferences suggesting little difference between control and PCOS conditions with mean point estimate of `r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day with a 95% quantile interval ranging from `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)` kcal/day. This was also the case for the standard deviation of REE with this model resulting in a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(main_arm_variance_effects_contrast_condition_bmi$estimate),2)` and a 95% quantile interval ranging from `r round(exp(main_arm_variance_effects_contrast_condition_bmi$conf.low),2)` to `r round(exp(main_arm_variance_effects_contrast_condition_bmi$conf.high),2)`. This was similar for fat-free mass adjusted models too showing little difference in mean REE between control and PCOS conditions with mean point estimate of `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day with a 95% quantile interval ranging from `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)` kcal/day and a contrast ratio between the control and PCOS conditions with a mean point estimate of `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$estimate),2)` and a 95% quantile interval ranging from `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$conf.low),2)` to `r round(exp(main_arm_variance_effects_contrast_condition_fat_free_mass$conf.high),2)`.

@@

This study sought to estimate and describe the magnitude of difference in REE between women with and without PCOS. Most studies identified in the systematic review, and included in the meta-analysis, used indirect calorimetry as the primary measure of REE and assessed women with PCOS who were insulin resistant and categorised as being in overweight or obese BMI categories compared to healthy controls. Our results indicate there is only a small magnitude of difference in REE (`r round(main_arm_mean_effects_contrast_condition$draw)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition$.lower)` kcal/day to `r round(main_arm_mean_effects_contrast_condition$.upper)`kcal/day]) between women with PCOS and those without. Further, there is little difference in between person variation between the groups based on the ratio of standard deviations (`r round(main_arm_variance_effects_contrast_condition$draw,2)` [95%QI: `r round(main_arm_variance_effects_contrast_condition$.lower,2)` to `r round(main_arm_variance_effects_contrast_condition$.upper,2)`]) suggesting that, despite individual differences in REE, PCOS is not systematically associated with lesser or greater individual variability.

These findings challenge the widely held belief that PCOS is inherently associated with a slower metabolism (PMID: 32244780), predisposing women with PCOS to weight gain. This belief may mistakenly lead to recommendations centred on slightly more severe calorie restriction to achieve weight-loss goals, compared with recommendations for the general population, as primary management strategies for women with PCOS [@ozgensaydamWeightManagementStrategies2021]. Recognising that there may be minimal differences in REE between women with PCOS and those without can inform both clinical and public practices, potentially leading to a shift in focus away from a requirement for more severe caloric restriction as a primary method of treatment towards more comprehensive, individualised, and psychologically safe approaches to care [@teedeRecommendations2023International2023].

The current study suggests that REE may not a barrier to weight regulation in PCOS given small group-level differences between women with PCOS and those without (`r round(main_arm_variance_effects_preds_condition$draw[2])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[2])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[2])` kcal/day] versus `r round(main_arm_variance_effects_preds_condition$draw[1])` kcal/day [95%QI:`r round(main_arm_variance_effects_preds_condition$.lower[1])` kcal/day to `r round(main_arm_variance_effects_preds_condition$.upper[1])` kcal/day], respectively). If anything, REE may be slightly higher in women with PCOS compared to healthy women without PCOS. BMI of women with PCOS ranged from `r summary_descriptives$min_bmi[2]` to `r summary_descriptives$max_bmi[2]` was typically greater than women without PCOS which ranged `r summary_descriptives$min_bmi[1]` to `r summary_descriptives$max_bmi[1]` and this may explain the slightly greater REE in the former group. However, for those studies where we could extract or estimate BMI, our additional exploratory models adjusted for this similarly showed little difference in REE (`r round(main_arm_mean_effects_contrast_condition_bmi$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_bmi$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_bmi$conf.high)`kcal/day]) between women with PCOS and those without. However, one of the studies included in our analysis [@tosiRestingEnergyExpenditure2024] reported that, whilst there was little difference in unadjusted REE, when REE was adjusted for fat-free mass it was lower in women with PCOS suggesting the potential importance of fat-free tissue in energy regulation. Yet, for those studies where fat-free mass was reported, our additional exploratory models adjusted for this similarly showed little difference in REE (`r round(main_arm_mean_effects_contrast_condition_fat_free_mass$estimate)` kcal/day [95%QI: `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.low)` kcal/day to `r round(main_arm_mean_effects_contrast_condition_fat_free_mass$conf.high)`kcal/day]) between women with PCOS and those without. As such, even adjusted for both BMI and fat-free mass, there seems to be little difference in REE between women with, and without, PCOS. Yet, a recent systematic review of mechanisms for metabolic dysfunction has reported excess androgen drives metabolic issues within adipose tissue and muscle tissue contributing to complications like obesity and insulin resistance (PMID: 32244780). Taken together, these findings highlight that factors beyond REE and typical correlates of this including BMI or fat-free mass, such as hormonal and tissue-specific metabolic effects, may play a more significant role in weight regulation challenges in women with PCOS.
